---
title: "Getting started with cWise"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with cWise}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{cWise}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(cWise)
```

## The crosswise model

Crosswise questions let respondents answer a sensitive question without
revealing their answer directly. A respondent is shown a sensitive statement
and a second statement whose prevalence is known (`p`). They report only
whether their answers to the two statements are the same or different. The
crosswise response is therefore not itself the sensitive-trait indicator.

The design protects privacy, but it also makes inattentive responses important:
respondents who answer at random pull the observed crosswise proportion toward
one half. `cWise` uses an *anchor* crosswise question with known sensitive-item
prevalence to estimate attentiveness and correct that bias.

The package includes simulated data from this design. `Y` is the primary
crosswise response, `A` is the anchor response, and `p` and `p.prime` are the
known randomization probabilities.

```{r data}
head(cmdata)
```

## Estimate prevalence

Use `bc_est()` to obtain both the naive and bias-corrected prevalence estimates.
Here, the anchor makes it possible to estimate the share of attentive
respondents as well.

```{r prevalence}
estimate <- bc_est(Y = Y, A = A, p = 0.15, p.prime = 0.15, data = cmdata)
estimate
```

`Results` contains point estimates, standard errors, and confidence intervals.
`Stats` reports the estimated attentive-response rate and the analysis sample
size. The bias-corrected row is usually the estimand of interest when the
anchor-question assumptions are credible.

## Sensitivity analysis without an anchor

When no anchor question was fielded, `cmBound()` shows how the estimate changes
over a plausible range of inattentive-response rates. The following calculation
uses only the observed primary-question proportion, its randomization
probability, and the sample size.

```{r bounds, fig.width = 7, fig.height = 4}
cmBound(lambda.hat = mean(cmdata$Y), p = 0.15, N = nrow(cmdata))
```

An optional direct-question estimate can be supplied with `dq` and `N.dq`; it
is then drawn as a reference line with its uncertainty interval.

## Simulation and power planning

Power-analysis functions are deliberately not evaluated while this vignette is
built: they run nested bootstrap simulations. Start with a small exploratory
run and increase `N.sim` for a study-planning analysis.

```{r power, eval = FALSE}
simulation <- sim_cwdata(
  N.sim = 200, sample = 500, prevalence = 0.10, p = 0.15, p.prime = 0.15,
  gamma = 0.80, direct = 0.05
)
simulation$Results

sim_power(
  N.sim = 500, sample = 1000, pi.null = 0.05, pi.alt = 0.10,
  p = 0.15, p.prime = 0.15, gamma = 0.80, direct = 0.05
)
```

`sim_estimates()` plots the simulated estimates, while `sim_power_N()` compares
the package's standard sample-size grid. These functions are intended for
interactive study planning rather than a package build.
